Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-22 writing/embedded_meta_learning_appendix.tex
Clarity, clarity, clarity

\appendix
The supplemental material is organized as follows: In section \ref{app_why_meta_mapping} we discuss the value of meta-mappings in detail by comparing to other methods of performing a new task. In section \ref{app_cards_tsne} we show $t$-sne results for the cards domain. In section \ref{app_supp_figures} we provide supplemental figures. In section \ref{app_lesion_results} we provide some lesion studies. In section \ref{app_detailed_methods} we list details fo the datasets and architectures we used, as well as providing links to the source code for all models, experiments, and analyses. In section \ref{app_numerical_results} we provide means and bootstrap CIs corresponding to the major figures in the paper. \par

\section{Why meta-map from tasks to tasks?} \label{app_why_meta_mapping}
Why are meta-mappings between tasks useful? To answer this, we consider various ways of adapting to a new task in figure \ref{fig_why_meta_mapping} (based on results from the cards domain, section \ref{sec_cards}). The system could adapt from seeing examples of the new task, but this requires going out and collecting data, which may be expensive and does not allow zero-shot adaptation. Alternatively, the system could perform the new task via a meta-mapping from a prior learned task, where the meta-mapping is either induced from examples of the meta-mapping, or from language. Finally, the system could perform a new task from language alone, if it is trained to map instructions to tasks. \par
To address this latter possibility, we trained a version of the model where we included training the language system to produce embeddings for the basic tasks (while simultaneously training the system on all the other objectives, such as performing the tasks from examples, in order to provide the strongest possible structuring of the system's knowledge for the strongest possible comparison). We compare this model's performance at held-out tasks to that of systems learning from examples of the new task directly, or from meta-mapping, see fig. \ref{fig_why_meta_mapping}. \par 
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/why_meta_mapping.png}
\caption{Comparison of a variety of methods for performing one of the 10\% held-out tasks in the more difficult hold-out set in the cards domain. There are a number of ways the system could adapt to a new task: from seeing example of the new task, from hearing the new task described from language alone, or from leveraging its knowledge about prior tasks via meta-mappings (in this case, from the non-losers variations of the same games). The meta-mappings offer a happy medium between the other two alternatives -- they only require cached knowledge of prior tasks, rather than needing to collect experience on the task before a policy can be derived, but they outperform a system that simply tries to construct the task embedding from a description alone. Language alone is not nearly as rich a cue as knowledge of how a new task relates to prior tasks.} 
\label{fig_why_meta_mapping}
\end{figure}
These results demonstrate the advantage of meta-mapping. While learning from examples is still better given enough data, it requires potentially-expensive data collection and does not allow zero-shot adaptation. Performing the new task from a language description alone uses only the implicit knowledge in the model's weights, and likely because of this it does not generalize well to the difficult held-out tasks. Meta-mapping performs substantially better, while relying only on cached prior knowledge, viz. prior task-embedding(s) and a description of the meta-mapping (either in the form of examples or natural language). That is, meta-mapping has the advantage of requiring no new data collection, like performing from language alone, but results in much better performance by leveraging a richer description of the new task constructed using the system's knowledge of a prior task and the new task's relationship to it.\par

>>>>

\appendix
The supplemental material is organized as follows: In section \ref{app_clarifying_meta_mapping} we clarify some definitional details, and discuss the value of meta-mappings in detail by comparing to other methods of performing a new task. In section \ref{app_cards_tsne} we show $t$-sne results for the cards domain. In section \ref{app_supp_figures} we provide supplemental figures. In section \ref{app_lesion_results} we provide some lesion studies. In section \ref{app_detailed_methods} we list details fo the datasets and architectures we used, as well as providing links to the source code for all models, experiments, and analyses. In section \ref{app_numerical_results} we provide means and bootstrap CIs corresponding to the major figures in the paper. \par

\section{Clarifying meta-mappings} \label{app_clarifying_meta_mapping}
\subsection{Why meta-map from tasks to tasks?} \label{app_why_meta_mapping}
Why are meta-mappings between tasks useful? To answer this, we consider various ways of adapting to a new task in figure \ref{fig_why_meta_mapping} (based on results from the cards domain, section \ref{sec_cards}). The system could adapt from seeing examples of the new task, but this requires going out and collecting data, which may be expensive and does not allow zero-shot adaptation. Alternatively, the system could perform the new task via a meta-mapping from a prior learned task, where the meta-mapping is either induced from examples of the meta-mapping, or from language. Finally, the system could perform a new task from language alone, if it is trained to map instructions to tasks. \par
To address this latter possibility, we trained a version of the model where we included training the language system to produce embeddings for the basic tasks (while simultaneously training the system on all the other objectives, such as performing the tasks from examples, in order to provide the strongest possible structuring of the system's knowledge for the strongest possible comparison). We compare this model's performance at held-out tasks to that of systems learning from examples of the new task directly, or from meta-mapping, see fig. \ref{fig_why_meta_mapping}. \par 
These results demonstrate the advantage of meta-mapping. While learning from examples is still better given enough data, it requires potentially-expensive data collection and does not allow zero-shot adaptation. Performing the new task from a language description alone uses only the implicit knowledge in the model's weights, and likely because of this it does not generalize well to the difficult held-out tasks. Meta-mapping performs substantially better, while relying only on cached prior knowledge, viz. prior task-embedding(s) and a description of the meta-mapping (either in the form of examples or natural language). That is, meta-mapping has the advantage of requiring no new data collection, like performing from language alone, but results in much better performance by leveraging a richer description of the new task constructed using the system's knowledge of a prior task and the new task's relationship to it.\par 
\subsection{A definitional note}
When we discussed meta-mappings in the main task, we equivocated between tasks and behaviors for the sake of brevity. For a perfect model, this is somewhat justifiable, because each task will have a corresponding optimal behavior, and the sytem's embedding of the task will be precisely the embedding which produces this optimal behavior. However, behavior-irrelevant details of the task, like the color of the board, may not be embedded, so this should not really be thought of as a task-to-task mapping. This problem is exacerbated when the system is imperfect, e.g. during learning. It is thus more precise to distinguish between a ground-truth meta-mapping, which maps tasks to tasks, and the computational approach to achieving that meta-mapping, which really maps between representations which combine both task and behavior. \par
\begin{figure}[H]
\includegraphics[width=\textwidth]{figures/why_meta_mapping.png}
\caption{Comparison of a variety of methods for performing one of the 10\% held-out tasks in the more difficult hold-out set in the cards domain. There are a number of ways the system could adapt to a new task: from seeing example of the new task, from hearing the new task described from language alone, or from leveraging its knowledge about prior tasks via meta-mappings (in this case, from the non-losers variations of the same games). The meta-mappings offer a happy medium between the other two alternatives -- they only require cached knowledge of prior tasks, rather than needing to collect experience on the task before a policy can be derived, but they outperform a system that simply tries to construct the task embedding from a description alone. Language alone is not nearly as rich a cue as knowledge of how a new task relates to prior tasks.} 
\label{fig_why_meta_mapping}
\end{figure}

### 2019-05-22 writing/embedded_meta_learning_main.tex
Clarity, clarity, clarity

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are in the form of meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

@@

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, tasks which take tasks as an input or output. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these meta-mappings are a useful way to think about human-like flexibility. For example, a go-playing system should be able to adapt to a different board size. Go on a large board and go on a small board share a great deal of structure. Humans can exploit this to play on a different board without retraining, but deep learning models generally have no way of achieving this. Similarly, suppose an agent knows how to navigate a town. It should be able to adapt to changes of the environment such as street closures. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not capture the extent of human flexibility. Meta-mapping allows us to adapt to task alterations zero-shot, that is, without seeing a single example from the new task. (See appendix \ref{app_why_meta_mapping} for a discussion of why this is better than e.g. conditioning on a task description.)\par
Meta-mappings can be cued in different ways, for example by seeing examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board''). Achieving this kind of flexibility will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
EML architectures are composed of two components: 
\vspace{-0.75em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explain behavior), or tasks to tasks (change behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

A basic forward pass through the system would proceed as follows. First, a training dataset is encoded by $\mathcal{I}$ and $\mathcal{T}$. These examples are fed to $\mathcal{M}$ to produce a function embedding. This function embedding is mapped through $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a second dataset, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on targets for the second dataset. See appendix \ref{app_model_details} for detailed architecture, operation, and hyper-parameters. \par 

@@

However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. For example, suppose we have some task embedding $z_{old} \in Z$, and a meta-mapping embedding $z_{\text{meta}} \in Z$, e.g. from $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{source,0},z_{target,0}, ...\right) \right\}\right)$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  
\[\hat{z}_{new} = f_{z_{\text{function}}}(z_{old}) \qquad \text{where } f_{z_{\text{meta}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]
For example, if the $z_{old}$ is the system's embedding for trying to win at go, and the meta-dataset consists of examples of a winning strategy and a losing strategy for different games, then $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate how well the system actually loses with the $\hat{z}_{new}$ strategy. All meta-mapping results reported here were evaluated in that way -- evaluating the loss on the target task after transforming the function embedding. \par

@@

\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The plots show the loss produced when evaluating the mapped embedding on the target task. For example, if the initial polynomial is $p(x) = x + 1$, and the meta-task is ``square,'' the loss would be evaluated by taking the embedding produced by applying this meta-mapping to the embedding of $p(x)$ and evaluating how well the mapped embedding regresses $p(x)^2 = x^2 + 2x + 1$. The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 

@@

\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results}. Furthermore, it is able to do this in a sample-efficient manner, see supp. fig. \ref{supp_poly_sweep_results}. \par 

>>>>

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks zero-shot. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. EML architectures are applicable to any type of machine learning task, including supervised learning and reinforcement learning. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior to adapt to new tasks. 
\end{abstract}

@@

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, tasks which take tasks as an input or output. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these task-to-task meta-mappings are a useful way to think about human-like flexibility. For example, a go-playing system should be able to adapt to a different board size. Go on a large board and go on a small board share a great deal of structure. Humans can exploit this to play on a different board without retraining, but deep learning models generally have no way of achieving this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture this human flexibility. We can adapt to task alterations zero-shot, that is, without seeing a single example from the new task. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
Their are different ways this adaptation can be cued, for example by seeing examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a discussion of why this is better than e.g. conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explain behavior), or meta-map tasks to tasks (change behavior). This is a step closer to the human flexibility. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\vspace{-0.5em}
\paragraph{Basic operation:} To perform a basic task, first, a training dataset is encoded by $\mathcal{I}$ and $\mathcal{T}$. These examples are fed to $\mathcal{M}$ to produce a function embedding. This function embedding is mapped through $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a second dataset, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on targets for the second dataset. See appendix \ref{app_model_details} for detailed architecture, operation, and hyper-parameters. \par 

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings. For example, suppose we have some embedding $z_{go} \in Z$ for the task of playing go. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of winning and trying-to-lose: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{checkers},z_{checkers,lose}\right), ... \right\}\right)$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  
\[\hat{z}_{new} = f_{z_{\text{function}}}(z_{go}) \qquad \text{where } f_{z_{\text{meta}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]
Then $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate whether the system loses with the $\hat{z}_{new}$ strategy. All meta-mapping results reported here were evaluated in that way -- evaluating the loss on the target task after transforming the task embedding. \par

@@

\caption{The system generalizes to apply learned meta-mappings to new polynomials, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The plots show the loss produced when evaluating the mapped embedding on the target task. For example, if the initial polynomial is $p(x) = x + 1$, and the meta-task is ``square,'' the loss would be evaluated by applying this meta-mapping to the embedding of $p(x)$ and evaluating how well the function specified by the mapped embedding regresses $p(x)^2 = x^2 + 2x + 1$. The solid line indicates optimal performance; the dashed line is chance (untrained). Error-bars are bootstrap 95\%-CIs.} 

@@

\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results} (with good sample-efficiency, see supp. fig. \ref{supp_poly_sweep_results}). \par 
